---
title: Realestate value in the Cape Peninsula
author: Zackary Pryde
date: '2022-01-17'
slug: []
categories: []
tags:
  - Geospatial
  - GIS
  - R
  - Web-scraping
---

## About this project

Cape Town, South Africa, is world renowned as beautiful landscape, an attribute that plays a dominant role in the real estate of this region. Property value in the mother city is generally a hot topic among locals and internationals. The general trends in property value over the region are typically known to some extent, and individuals that are interested often map this out by looking over various listings on the web. Here, I aim to summarize property value at the suburban level over the Southern region of the City of Cape Town Metropolitan Municipality (i.e., The Cape Peninsula and surroundings).

Using the R programming language, I write some basic web-scraping code to obtain property value data from a popular website - [Property24](https://www.property24.com/about-us) - over a specified [set of filters](https://www.property24.com/for-sale/advanced-search/results?sp=cid%3d615%2c479%2c478%2c652%2c475%26s%3d11742%2c9044%2c15183%2c9057%2c9040%2c9034%2c9025%2c9047%2c10211%2c12788%2c16635%2c10180%2c9036%2c10213%2c10174%2c9039%2c9067%2c9069%2c9097%2c10178%2c11740%2c12840%2c15132%2c16720%2c10094%2c10195%2c10052%2c10189%2c10207%2c10096%2c10170%2c10204%2c10212%2c10090%2c10179%2c10109%2c10102%2c8778%2c8792%2c8812%2c9118%2c11014%2c11012%2c8661%2c14225%2c14224%2c8669%2c11741%2c8682%2c8679%2c8806%2c8787%2c8734%2c8783%2c8779%2c10157%2c8017%2c9153%2c8683%2c8677%2c8686%2c8717%2c8754%2c9163%2c9145%2c9149%2c9155%2c9166%2c10166%2c9138%2c11021%2c11016%2c11013%2c11015%2c11017%2c9169%2c11018%2c9136%2c9141%2c9143%2c10164%2c10163%2c10161%2c16541%2c10158%2c9119%2c8731%2c8736%2c8730%2c8716%2c8800%2c9029%2c12195%2c17446%2c17447%2c15627%2c9103%2c9101%2c9106%2c15622%2c9104%2c16197%2c9095%2c7851%2c9110%2c9102%2c9105%2c9108%2c9085%2c9107%2c10603%2c9062%2c17448%2c9094%2c16008%2c10203%2c8025%2c8010%2c8005%2c8019%26so%3dNewest#SortOrder). As to be expected, web-scraping data may typically fall short in terms of it's quality. The data acquisition is hence followed by a collaboration of data-wrangling in R, and manipulation/cleaning using graphical GIS software, QGIS. 

Once the data are appropriately structured, I construct an interactive map that displays the median cost of purchasing a house in a given suburb of the study extent. 

## How?

### Step 1

### Step 2

### Step 3


## The result

```{r, include=FALSE}
# 1. Package installations ----

# install.packages("rvest")
# install.packages("dplyr")
# install.packages("tmap")
# install.packages("leaflet")
# install.packages("tmaptools")

# 2. Loading the required libraries ----

library(rvest)
library(dplyr)
library(tidyr)
library(rgdal)
library(tidyverse)
library(ggmap)
library(sf)
library(mapview)
library(tmap)
library(leaflet)
library(tmaptools)

```

```{r, include=FALSE}
CPM = st_read("CapePenMap.shp")
CPMn = st_as_sf(CPM)
CPMn = data.frame("Suburb" = CPM$NAME, "Median Price" = CPM$CP_Final2_, "No. listed" = CPM$CP_Final_1, CPM$geometry)
CPMn = st_as_sf(CPMn)
```

```{r, include=FALSE}
# P = ggplot(CPMn) + 
#   geom_sf(aes(fill = Median.Price)) + 
#   scale_fill_gradient2()
# 
# Nl = ggplot(CPMn) + 
#   geom_sf(aes(fill = No..listed)) + 
#   scale_fill_gradient2()
```

```{r, echo=FALSE,message=FALSE}
mapview(CPMn, zcol = "Median.Price")
```

## If you are interested...

You can find the full project [here](https://github.com/ZackaryPryde/CP_Realestae) on GitHub. 

